---
title: "Untitled"
author: "Brian Maitner"
date: "10/6/2022"
output: html_document
---

# Terrestrial sampling options

In this file, the goal is to:

  1) Examine the environmental coverage available within focal sites
  2) Attempt to subdivide those focal sites into clusters for representative sampling


```{r setup, include=FALSE, echo=FALSE}

# Load required packages

library(googledrive)
library(terra)
library(ggplot2)
library(stars)
library(sf)
library(movecost) #https://www.sciencedirect.com/science/article/pii/S2352711019302341#fig2
library(piggyback)
source("https://raw.githubusercontent.com/AdamWilsonLab/emma_envdata/main/R/robust_pb_download.R")


#Load required data

domain <- st_read("data/output/domain.gpkg") #the full domain
sampling_options <- st_read("data/output/sampling_options.gpkg") #parks falling within the domain
acceptable_sites <- st_read("data/output/acceptable_sites.gpkg")


```

# Sampling options

```{r, echo=FALSE}


ggplot()+
  geom_sf(data = domain, col = "black")+
  geom_sf(data = sampling_options,fill="dark grey",col = "dark grey")+
  geom_sf(data = acceptable_sites, fill="red", col = "red")


```

Figure 1. The full domain (light grey), available study sites (dark grey), and focal study sites (red).

# Focal sites

Focal study sites have been selected on the basis of the following criteria:

* occur in the domain
* occur within a national park
* occur within 1 km of a road
* have a mean NDVI over the last year of > 0.2
* have a slope < 30 degrees
* are within an estimated hiking time of 2 hours from the road

# Clustering

To better stratify our sampling, we'll use clustering to 


